---
title: "Coursera Project - Reproducible-research"
output: html_document
---

##Reading the dataset 

```{r}
knitr::opts_chunk$set(echo = TRUE)
suppressWarnings(library(lattice))
```

```{r}

if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
       temp <- tempfile()
      download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
      unlink(temp)
}
smpdata <- read.csv("activity.csv")

```

##Count number of steps each day, draw histogram, calculate mean and median
```{r}
perdaysteps <- aggregate(steps~date, data = smpdata, na.rm=TRUE, sum)

hist(perdaysteps$steps, breaks= 20, xlab = "Per Day Steps", main = "Total Steps per day")

smpmean <- mean(smpdata$steps, na.rm=TRUE)
print(paste("The mean step each day- ", smpmean))

smpmedian <- median(smpdata$steps, na.rm = TRUE)
print(paste("The median of step each day- ", smpmedian))
```

##Time series plot for steps at 5 minute interval and interval in which maximum steps taken.
```{r}
avgsteps <- aggregate(steps~interval, data = smpdata, na.rm=TRUE, mean)
plot(avgsteps$interval, avgsteps$steps, type="l", xlab="Interval", ylab = "Steps", main="Average steps in 5 minutes interval")

maxstpintv <- avgsteps[avgsteps$steps==max(avgsteps$steps),]
#print(paste("The 5 minute interval with max steps is -", maxstpintv))

```

##Total missing values in data set, mean and median after imputation of missing values
```{r}
print(paste("Total missing values - ",sum(is.na(smpdata$steps))))

avgdatestp <- aggregate(steps ~ date, data= smpdata, na.rm=TRUE, mean)

imputedata <- transform(smpdata, steps = ifelse(is.na(smpdata$steps), avgdatestp$steps[match(smpdata$date, avgdatestp$date)], smpdata$steps))
imputedata[as.character(imputedata$date) == "2012-10-01",1] <- 0

perdaysteps_new <- aggregate(steps ~ date, data = imputedata, sum)

hist(perdaysteps_new$steps, xlab = "Steps", main="Per day steps - Imputed")

smpmean_new <- mean(imputedata$steps, na.rm=TRUE)
print(paste("The mean step each day- ", smpmean_new))

smpmedian <- median(smpdata$steps, na.rm = TRUE)
print(paste("The median of step each day- ", smpmedian))

```

The mean of the imputed data has reduced as compared to the data with NA's. But the median is same as 0. 

##The average number of steps taken, averaged across all weekday days or weekend days 

```{r}
smpdata$Day <- weekdays(as.Date(smpdata$date))
smpdata$Day<- as.factor(ifelse(smpdata$Day == "Saturday" |smpdata$Day == "Sunday", "weekend", "weekday"))

avgstepsweekday <- aggregate(steps~interval + Day, data = smpdata, na.rm=TRUE, mean)

xyplot(avgstepsweekday$steps~avgstepsweekday$interval |avgstepsweekday$Day , layout=c(1,2), type="l", xlab="Interval", ylab ="Number of Steps", main="Average steps in interval" )

```